Variation in distance between consecutive animal locations

### 2021-03-15 distributions_paper/rudolph_distributions.Rmd
Building distribution fits

  dist.used <- c("lnorm", "weibull", "cauchy")
  param <- c("meanlog", "sdlog", "shape", "scale", "location", "scale")
  dist <- c("lnorm", "lnorm", "weibull", "weibull", "cauchy", "cauchy")
}

@@

  dist.used <- c( "exp", "gamma", "weibull", "lnorm")
  param <- c("rate", "shape", "rate", "shape", "scale", "meanlog", "sdlog")
  dist <- c("exp", "gamma", "gamma", "weibull", "weibull", "lnorm", "lnorm")
}

@@

```{r eval=FALSE}
descdist(df$dist, boot = 1000)
hist(df$dist)

# problem
fitdist(df$dist, distr = "exp")
fitdist(df$dist, distr = "gamma")
#no issue
fitdist(df$dist, distr = "lnorm")
fitdist(df$dist, distr = "weibull")
fitdist(df$dist, distr = "cauchy")

@@

prms.df <- build_fits_df(pop) %>% 
  mutate(param = param,
         dist = dist,
         data = "pop")

@@

# Fitting the four distributions for each individual and saving the parameters in a dataframe
for(i in 1:length(inds)){
  fit.df <- df %>% 
    dplyr::filter(., Bird_ID == inds[i])

@@

@@

  fit.prms <- build_fits_df(fit) %>% 
  mutate(param = param,
         dist = dist,
         data = inds[i])

@@

# Getting qqplot data for each individual
inds.models.data <- NULL
ind.models <- list(B1, B13, B19, B22, B28, B29, B3, B30, B49, B5, B7, B84)
# ind.models <- list(B1, B13, B17, B19, B20, B22, B24, B28, B29, B3, B30, B49, B5, B7, B84)
chr.models <- inds
for(i in 1:length(chr.models)){
  x <- ind.models[[i]]

  qqdata <- qqcomp(x, plotstyle = "ggplot")$data %>% 
    mutate(data = chr.models[i])
  names(qqdata) <- c("theoretical", "fdist", "empirical", "individual")
  inds.models.data <- bind_rows(inds.models.data, qqdata)
}
```

@@

>>>>

  dist.used <- c("lnorm", "weibull", "cauchy")
  param <- c("meanlog", "sdlog", "shape", "scale", "location", "scale")
  dist <- c("lnorm", "lnorm", "weibull", "weibull", "cauchy", "cauchy")
  kpars <- c(2, 2, 2, 2, 2, 2)
}

@@

  dist.used <- c( "exp", "gamma", "weibull", "lnorm")
  param <- c("rate", "shape", "rate", "shape", "scale", "meanlog", "sdlog")
  dist <- c("exp", "gamma", "gamma", "weibull", "weibull", "lnorm", "lnorm")
  kpars <- c(1, 2, 2, 2, 2, 2, 2)
}

if(focus == "m15"){
  df <- aracari_df %>% 
  dplyr::filter(dist != 0) %>% 
  group_by(id) %>%
  filter(., n() >= 30) %>%
  ungroup() %>%
  mutate(tmin = dt/60,
         mpm = dist/tmin,
         m15 = mpm*15,
         Bird_ID = paste0("B", id),
         focus = m15) %>% 
  dplyr::filter(., m15 != 0)

  dist.used <- c( "exp", "gamma", "weibull", "lnorm")
  param <- c("rate", "shape", "rate", "shape", "scale", "meanlog", "sdlog")
  dist <- c("exp", "gamma", "gamma", "weibull", "weibull", "lnorm", "lnorm")
  kpars <- c(1, 2, 2, 2, 2, 2, 2)
}

@@

@@

prms.df <- build_fits_df(pop) %>% 
  mutate(param = param,
         dist = dist,
         kpars = kpars,
         data = "pop")

@@

# Fitting the four distributions for each individual and saving the parameters in a dataframe
ind.models <- NULL
for(i in 1:length(inds)){
  fit.df <- df %>% 
    dplyr::filter(., Bird_ID == inds[i])

@@

  ind.models[[i]] <- fit

@@

  fit.prms <- build_fits_df(fit) %>% 
  mutate(param = param,
         dist = dist,
         kpars = kpars,
         data = inds[i])

@@

# Getting qqplot data for each individual
inds.models.data <- NULL
for(i in 1:length(inds)){
  x <- ind.models[[i]]

  qqdata <- qqcomp(x, plotstyle = "ggplot")$data %>% 
    mutate(data = inds[i])
  names(qqdata) <- c("theoretical", "fdist", "empirical", "individual")
  inds.models.data <- bind_rows(inds.models.data, qqdata)
}
```

@@

## Information criteria

ic_calc <- function(n, kpars, loglik){

  BIC <- log(n)*kpars - 2*loglik
  AIC <- 2*kpars - 2*loglik
  AICc <- AIC + (2* kpars * (kpars+1)/(n-kpars-1))
}

prms.df %>% 
  mutate(BIC = log(n)*kpars - 2*loglik,
         AIC = 2*kpars - 2*loglik,
         AICc = AIC + (2* kpars * (kpars+1)/(n-kpars-1))) %>% 
  group_by(data) %>% 
  mutate(deltaBIC = signif(BIC - min(BIC), 3),
         deltaAIC = signif(AIC - min(AIC), 3),
         deltaAICc = signif(AICc - min(AICc), 3)) -> prms_ic
```

So, for each individual, which is the best fitting distribution, based on AIC or BIC? Here we are just focusing on the model fits for each individual.

```{r}
prms_ic %>% 
  filter(., data != "pop") %>% 
  dplyr::select(dist, data, deltaAIC,
                #deltaAICc,
                deltaBIC) %>% 
  distinct() %>% 
  pivot_longer(., cols = starts_with("delta"), names_to = "ic", values_to = "value") %>% 
  filter(value == 0) %>% 
  dplyr::select(-value) %>% 
  pivot_wider(., names_from = ic, values_from = dist) %>% 
  mutate(agree = ifelse(deltaAIC == deltaBIC, "yes", "no"))
```

```{r fig.width=12, fig.height=4}
# To visualize the differences in the information criteria at the individual level

prms_ic %>% 
  filter(., data != "pop") %>% 
  dplyr::select(dist, data, deltaAIC, 
                deltaAICc,
                deltaBIC) %>% 
  distinct() %>% 
  pivot_longer(., cols = starts_with("delta"), names_to = "ic", values_to = "value") %>% 
  mutate(plot_alpha = ifelse(value == "0", 1, 0.8)) %>% 
  ggplot(., aes(x = data, y = value, color = dist 
                #alpha = plot_alpha
                )) +
  facet_wrap(~ic) +
  geom_point(size = 3) +
  scale_color_brewer(palette = "Spectral") + 
  #scale_alpha_continuous(range = c(0.5, 1)) +
  labs(x = "") +
  theme(legend.position = "bottom") +
  guides(color = guide_legend(title = "Distribution"))

```
Now we can compare what we call *models*, in which we compare a model that does complete pooling and considers all the data together, and the other model that considers individual variation. 
So, the complete pooling, takes all the data and fits a distribution to it. 
The individual variation models will fit a distribution to each individual. At one level we will use the same distribution for all individuals and only consider variation in parameters. At the next level, we will consider a change in parameters and distributions, where individuals can have different distributions, and this is what we call the mixed distribution model for individual variation.
